### 2016-09-27 syntax.tex

%Informally, the representation should serve as a good
%approximation of the multiset, at least with respect to the allowed queries.
\begin{definition}[Data structure] \rm
Fix sets~${\cal U}, \Gamma$, and $\calQ$, where $\calQ = \{\qry
\colon {\cal U} \to \Gamma\}$. A \emph{data structure} (supporting
queries $\calQ$ over universe ${\cal U}$) is a tuple
$\setprim=(\Rep, \Qry)$ where:
\begin{itemize}
\item $\Rep \colon {\cal U} \to \bits^* \times \bits^*$ is a
    randomized \emph{representation algorithm}, taking as input
    an element $S \in \mathcal{U}$,
    and outputting %a \emph{representation} $M \in \Sigma^*$,
    \emph{public representation data} $\pubaux \in \bits^*$ and
    \emph{private representation data} $\privaux \in \bits^*$.
    We write $(\pubaux,\privaux) \getsr \Rep(S)$ as shorthand
    for this execution.
\item $\Qry\colon \bits^* \times \bits^* \times \calQ \to
    \Gamma$ is a deterministic \emph{query-evaluation}
    algorithm, taking as input $\pubaux,\privaux \in \bits^*$
    and a query $\qry \in \calQ$, and outputting an
    \emph{answer} $a \in \Gamma$.  We write $a \gets
    \Qry(\pubaux,\privaux,\qry)$ for this execution. \hfill\dqed
\end{itemize}
\end{definition}

@@

\begin{definition}[Error probability and rate] \rm
For a data structure $\setprim=(\Rep, \Qry)$ supporting queries
${\cal Q}$ over universe ${\cal U}$, let
$\mathrm{Err}(S,\qry,\pubaux,\privaux)$ be  true iff
$\Qry(\pubaux,\privaux, \qry) \neq \qry(S)$.  Then $\Pi$ has
\emph{error probability} $\epsilon$ if
$\max_{S,\qry}\Pr[(\pubaux,\privaux) \getsr \Rep(S) :
\mathrm{Err}(S,\qry,\pubaux,\privaux)=1 ] \leq \epsilon$.
%
It has \emph{error rate} $\tilde{\epsilon}$ if
$\max_{S}\Pr[(\pubaux,\privaux) \getsr \Rep(S);\, \qry \getsr \calQ :
\mathrm{Err}(S,\qry,\pubaux,\privaux)=1] \leq \tilde{\epsilon}$. \hfill\dqed
\end{definition}

@@

These definitions, and those that follow, can be lifted to the random-oracle
model by giving $\Rep, \Qry$ access to a random oracle. In that
case, the probabilities defining the error probability and error
rate are also taken over choice of the random oracle.

We note that, while our error probability and error rate do not
distinguish among ``types'' of errors, they also do not preclude
instantiations (like the classical Bloom filter) that admit only one
type of error, e.g., false positives.

\def\bin{{\sf Bin}}
\heading{Error probability vs.\ error rate. }
We pause to highlight an important
difference between error probability and error rate.  In particular,
the error probability~$\epsilon$ is measured only over the coins
of~$\Rep$, for a fixed~$S$ and query~$\qry$. Once the coins of~$\Rep$
are fixed, it is not straightforward how to connect~$\epsilon$ to
events that depend on other sources of randomness.   We illustrate how
this surfaces with the following example.

\jnote{The following may need to be updated to fit the new syntax.}
Let~$\Pi$ be a data structure that has error-probability $\epsilon$,
and consider a correctness-violating attack that asks random queries
$\qry_1,\qry_2,\ldots,\qry_T$, i.e., uniform samples from~$\calQ$.
Let $\mathrm{Err}_i$ be a random variable indicating the event
$\Qry(\pubaux,\privaux,\qry_i)\neq \qry_i(S)$.  We know that the
probability that $\mathrm{Err}_i=1$ is at most~$\epsilon$, and it is
tempting to model the~$\mathrm{Err}_i$ as i.i.d. Bernoulli trials
with success probability~$\epsilon$. However, a moment's reflection
shows that this is not the case. Concretely, consider the following
scheme. Fix $|S|=n$ for all $S \in \calS$, where $n \ll |\univ|$.
For the queries, let $\calQ = \{\qry_x\colon \calS \to \bits \,|\,
\forall x \in \univ\}$ where $\forall S \in \calS$ the predicate
$\qry_x(S)=1 \Leftrightarrow x \in S$.  Let $\Rep(S)$ pick a random
element~$s \in S$, and return $M = S \setminus \{s\}$ and $\privaux
= \emptystring$.  To respond to queries,
let~$\Qry(\pubaux,\privaux,\qry_x)=1$ iff $x \in M$.  It is easy to
see that this scheme is $1/n$-correct; for any fixed~$S$, there are
exactly~$n$ queries~$\qry_x$ that have probability~$1/n$ (over the
coins of~$\Rep$) of causing a correctness error, and the remaining
$|\univ|-n$ queries have probability zero of causing an error.
Moreover, once the coins of~$\Rep$ are fixed, there is exactly
\emph{one} query that will cause an error.  Thus for random queries
from~$Q$, for all~$i\in[T]$ we have
$\Prob{\mathrm{Err}_i=1}=1/|\univ| \ll 1/n = \epsilon$.

\heading{Data structures for set-membership queries.} Our particular
focus will be on data structures that support membership queries.
\jnote{I changed this from multiplicity to membership.}

@@

\jnote{Clean up the following, because we discuss BFs later.}

The classical Bloom filter is a special case of a set-multiplicity
data structure, as set-membership queries are a type of
set-multiplicity query.  Specifically,
%\jnote{Only handles sets, not multisets.}
let $\Gamma=\bits$, and fix functions $k(\cdot)$ and $m(\cdot)$
parameterizing the scheme.  Algorithm $\Rep$, on input a set~$S$,
computes $k=k(|S|)$ and $m=m(|S|)$ and then chooses $k$ functions
$h_1, \ldots, h_k \in \mathcal{H}$ for some function family
$\mathcal{H}=\{h \colon \univ \to \mathbb{N}\}$.\footnote{For this
  example, we are implicitly assuming that the range of the hash
  functions admits a natural mapping to $[m]$.  In practice, this
  could be enforced in various ways, including allowing $\Rep$ to
  return an indication of error. }
It then computes an $m$-bit
array~$M$ (initialized to 0 everywhere) by setting $M[h_i(x)]=1$ for
all $i\in [k]$ and $x \in S$.  Finally, $\Rep$ returns
$\privaux=\emptystring$ and $\pubaux = \langle
M,h_1,h_2,\ldots,h_k\rangle$ as the representation.
%
Algorithm~$\Qry$ is defined so that, on input
$(\pubaux,\privaux,\qry_x)$, it returns the minimum of the values
$M[h_1(x)],\ldots,M[h_k(x)]$.  We note that the standard Bloom filter
analysis treats $h_1,h_2,\ldots,h_k$ as independent random functions,
which we may capture in the ROM with some simple adjustments to the above.

@@

\tsnote{needs updating, since $\Sigma$ is gone, and the range of
  $\Rep$ is just $\bits^*$ with implicit encoding of whatever is the
  ``natural'' way to think about $\pubaux$ (e.g. counters, xor-shares)}
By setting $\Sigma=\bits^L$ for a specified $L>0$, we can capture
the so-called ``garbled Bloom filter'' due to Dong, Chen and
Wen~\cite{dong2013private}. \todo{Details have to be filled in.} This stores
xor-shares of bitstring~$x \in S$ at positions
$h_1(x),h_2(x),\ldots,h_k(x)$ in an array. Similarly, the
cuckoo-hashing construction from NY, which (loosely speaking) stores
a hash $g(x)$ at positions determined by
$h_1(x),h_2(x),\ldots,h_k(x)$, is captured when $L=|g(x)|$.
\todo{Check the correctness of this claim.}

>>>>

%Informally, the representation should serve as a good
%approximation of the multiset, at least with respect to the allowed queries.
\begin{definition}[Data structure] \rm
Fix sets~${\cal U}, \Gamma$, and $\calQ = \{\qry \colon {\cal U} \to
\Gamma\}$. A \emph{data structure} (supporting queries $\calQ$ over
universe ${\cal U}$) is a tuple $\setprim=(\Rep, \Qry)$ where:
\begin{itemize}
\item $\Rep \colon {\cal U} \to \bits^* \times \bits^*$ is a
    randomized \emph{representation algorithm}, taking as input
    an element $S \in \mathcal{U}$,
    and outputting %a \emph{representation} $M \in \Sigma^*$,
    \emph{public representation data} $\pubaux \in \bits^*$ and
    \emph{private representation data} $\privaux \in \bits^*$.
    We write $(\pubaux,\privaux) \getsr \Rep(S)$ as shorthand
    for this execution.
\item $\Qry\colon \bits^* \times \bits^* \times \calQ \to
    \Gamma$ is a deterministic \emph{query-evaluation}
    algorithm, taking as input $\pubaux,\privaux \in \bits^*$
    and a query $\qry \in \calQ$, and outputting an
    \emph{answer} $a \in \Gamma$.  We write $a \gets
    \Qry(\pubaux,\privaux,\qry)$ for this execution. \hfill\dqed
\end{itemize}
\end{definition}

@@

\begin{definition}[Error probability and rate] \rm
For a specific data structure $\setprim=(\Rep, \Qry)$ supporting
queries ${\cal Q}$ over universe ${\cal U}$, let
$\mathrm{Err}(S,\qry,\pubaux,\privaux)$ be  true iff
$\Qry(\pubaux,\privaux, \qry) \neq \qry(S)$.  Then we say that $\Pi$
has \emph{error probability}~$\epsilon$ if
\[\max_{S,\qry}\Pr[(\pubaux,\privaux) \getsr \Rep(S) :
\mathrm{Err}(S,\qry,\pubaux,\privaux)=1 ] \leq \epsilon.\]
%
We say it has \emph{error rate} $\tilde{\epsilon}$ if
\[{\textstyle \max_{S}\Pr[(\pubaux,\privaux) \getsr \Rep(S);\, \qry \getsr \calQ :
\mathrm{Err}(S,\qry,\pubaux,\privaux)=1]} \leq \tilde{\epsilon}. \]
%\hfill\dqed
\end{definition}

@@

The above definition can be lifted to the random-oracle model by
giving $\Rep, \Qry$ access to a random oracle. In that case, the
probabilities are also taken over choice of the random oracle. We
remark that, while our error probability and error rate do not
distinguish among ``types'' of errors, they also do not preclude
instantiations (like the classical Bloom filter) that admit only one
type of error, e.g., false positives.

\def\bin{{\sf Bin}}
\ignore{
\heading{Error probability vs.\ error rate. } We pause to
highlight an important difference between error probability and
error rate.  In particular, the error probability~$\epsilon$ is
measured only over the coins of~$\Rep$, for a fixed~$S$ and
query~$\qry$. Once the coins of~$\Rep$ are fixed, it is not
straightforward how to connect~$\epsilon$ to events that depend on
other sources of randomness.   We illustrate how this surfaces with
the following example.

\jnote{The following may need to be updated to fit the new syntax.}
Let~$\Pi$ be a data structure that has error-probability $\epsilon$,
and consider a correctness-violating attack that asks random queries
$\qry_1,\qry_2,\ldots,\qry_T$, i.e., uniform samples from~$\calQ$.
Let $\mathrm{Err}_i$ be a random variable indicating the event
$\Qry(\pubaux,\privaux,\qry_i)\neq \qry_i(S)$.  We know that the
probability that $\mathrm{Err}_i=1$ is at most~$\epsilon$, and it is
tempting to model the~$\mathrm{Err}_i$ as i.i.d. Bernoulli trials
with success probability~$\epsilon$. However, a moment's reflection
shows that this is not the case. Concretely, consider the following
scheme. Fix $|S|=n$ for all $S \in \calS$, where $n \ll |\univ|$.
For the queries, let $\calQ = \{\qry_x\colon \calS \to \bits \,|\,
\forall x \in \univ\}$ where $\forall S \in \calS$ the predicate
$\qry_x(S)=1 \Leftrightarrow x \in S$.  Let $\Rep(S)$ pick a random
element~$s \in S$, and return $M = S \setminus \{s\}$ and $\privaux
= \emptystring$.  To respond to queries,
let~$\Qry(\pubaux,\privaux,\qry_x)=1$ iff $x \in M$.  It is easy to
see that this scheme is $1/n$-correct; for any fixed~$S$, there are
exactly~$n$ queries~$\qry_x$ that have probability~$1/n$ (over the
coins of~$\Rep$) of causing a correctness error, and the remaining
$|\univ|-n$ queries have probability zero of causing an error.
Moreover, once the coins of~$\Rep$ are fixed, there is exactly
\emph{one} query that will cause an error.  Thus for random queries
from~$Q$, for all~$i\in[T]$ we have
$\Prob{\mathrm{Err}_i=1}=1/|\univ| \ll 1/n = \epsilon$.}

\heading{Data structures for set-membership queries.} An important
special case is a data structure that supports set-membership
queries. \jnote{I changed this from multiplicity to membership.}

@@

Note that we do not require ${\cal U} = 2^\univ$ since, e.g., a
set-membership data structure may only be defined for sets up to
some bounded size.

\ignore{ \jnote{Below needs to change.} For concreteness, we note
that a classical Bloom filter is a particular example of a
set-membership data structure. Specifically,
%\jnote{Only handles sets, not multisets.}
let $\Gamma=\bits$, and fix functions $k(\cdot)$ and $m(\cdot)$
parameterizing the scheme.  Algorithm $\Rep$, on input a set~$S$,
computes $k=k(|S|)$ and $m=m(|S|)$ and then chooses $k$ functions
$h_1, \ldots, h_k \in \mathcal{H}$ for some function family
$\mathcal{H}=\{h \colon \univ \to \mathbb{N}\}$.\footnote{For this
  example, we are implicitly assuming that the range of the hash
  functions admits a natural mapping to $[m]$.  In practice, this
  could be enforced in various ways, including allowing $\Rep$ to
  return an indication of error. }
It then computes an $m$-bit
array~$M$ (initialized to 0 everywhere) by setting $M[h_i(x)]=1$ for
all $i\in [k]$ and $x \in S$.  Finally, $\Rep$ returns
$\privaux=\emptystring$ and $\pubaux = \langle
M,h_1,h_2,\ldots,h_k\rangle$ as the representation.
%
Algorithm~$\Qry$ is defined so that, on input
$(\pubaux,\privaux,\qry_x)$, it returns the minimum of the values
$M[h_1(x)],\ldots,M[h_k(x)]$.  We note that the standard Bloom filter
analysis treats $h_1,h_2,\ldots,h_k$ as independent random functions,
which we may capture in the ROM with some simple adjustments to the above.

@@

\tsnote{needs updating, since $\Sigma$ is gone, and the range of
  $\Rep$ is just $\bits^*$ with implicit encoding of whatever is the
  ``natural'' way to think about $\pubaux$ (e.g. counters, xor-shares)}
By setting $\Sigma=\bits^L$ for a specified $L>0$, we can capture
the so-called ``garbled Bloom filter'' due to Dong, Chen and
Wen~\cite{dong2013private}. \todo{Details have to be filled in.}
This stores xor-shares of bitstring~$x \in S$ at positions
$h_1(x),h_2(x),\ldots,h_k(x)$ in an array. Similarly, the
cuckoo-hashing construction from NY, which (loosely speaking) stores
a hash $g(x)$ at positions determined by
$h_1(x),h_2(x),\ldots,h_k(x)$, is captured when $L=|g(x)|$.
\todo{Check the correctness of this claim.} }
